The Proofs

### 2023-11-13 proofs/tex/au-lemmas.tex
Restructure proofs

\begin{observation}[Anti-unification algorithm is deterministic]
    \label{obs:au-deterministic}
    \hfill\\
\begin{itemize}
    \item [$+$] If $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$ and 
    $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ', uQ', aus'1, aus'2)]]$,
    then $[[Ξ]] = [[Ξ']]$, $[[uQ]] = [[uQ']]$, $[[aus1]] = [[aus'1]]$, and $[[aus2]] = [[aus'2]]$.
    \item [$-$] If $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$ and
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ', uM', aus'1, aus'2)]]$,
    then $[[Ξ]] = [[Ξ']]$, $[[uM]] = [[uM']]$, $[[aus1]] = [[aus'1]]$, and $[[aus2]] = [[aus'2]]$.
\end{itemize}
\end{observation}
\begin{proof}
    By trivial induction on $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
    and mutually on $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.
\end{proof}

\begin{observation} \label{obs:names-defined-by-mapping}
    Names of the anti-unification variables are uniquely defined by
    the types they are mapped to by the resulting substitutions.

    \begin{itemize}
        \item [$+$]  Assuming $[[iP1]]$ and $[[iP2]]$ are normalized,
        if $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
        then for any $[[β̂⁻]] \in [[Ξ]]$,
        $[[β̂⁻]] = [[â⁻_{[aus1]β̂⁻, [aus2]β̂⁻}]]$
        \item [$-$]  Assuming $[[iN1]]$ and $[[iN2]]$ are normalized,
        if $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
        then for any $[[β̂⁻]] \in [[Ξ]]$,
        $[[β̂⁻]] = [[â⁻_{[aus1]β̂⁻, [aus2]β̂⁻}]]$
    \end{itemize}
\end{observation}
\begin{proof}
    By simple induction on $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
    and mutually on $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.
    Let us consider tha last rule applied to infer this judgment.
    \begin{caseof}
        \item \ruleref{\ottdruleAUPVarLabel} or \ruleref{\ottdruleAUNVarLabel},
        then $[[Ξ]] = [[·]]$, and the property holds vacuously.

@@

\begin{lemma}[Soundness of Anti-Unification] \label{lemma:au-soundness}
    \hfill
    \begin{itemize}
        \item [$+$]  Assuming $[[iP1]]$ and $[[iP2]]$ are normalized,
        if $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
        then 
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uQ]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification substitutions, and
            \item $[[ [ausi] uQ = iPi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}

        \item [$-$] Assuming $[[iN1]]$ and $[[iN2]]$ are normalized,
        if $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
        then
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uM]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification substitutions, and
            \item $[[ [ausi] uM = iNi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on 
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
    and mutually, $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$.
    Let us consider the last rule applied to infer this judgement.
    \begin{caseof}
        \item \ruleref{\ottdruleAUNVarLabel}, then $[[iN1]] = [[α⁻]] = [[iN2]]$,
              $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, and $[[aus1]] = [[aus2]] = [[·]]$.
            \begin{enumerate}
                \item $[[Γ ; · ⊢ α⁻]]$ follows from the assumption $[[Γ ⊢ α⁻]]$,
                \item $[[Γ ; · ⊢ · : ·]]$ holds trivially, and
                \item $[[ [·] α⁻ = α⁻ ]]$ holds trivially.
            \end{enumerate}
        \item \label{case:anti-unification-soundness:shiftu}
         \ruleref{\ottdruleAUShiftULabel}, then $[[iN1]] = [[↑iP1]]$,
                $[[iN2]] = [[↑iP2]]$, and the algorithm makes the recursive call:
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
                returning $[[(Ξ, ↑uQ, aus1, aus2)]]$ as the result.

@@

% \renewcommand{\ottdruleAUPVarName}[0]{(Var$^{+[[≈au]]}$)}
% \renewcommand{\ottdruleAUShiftDName}[0]{($\downarrow^{[[≈au]]}$)}
% \renewcommand{\ottdruleAUExistsName}[0]{($\exists^{[[≈au]]}$)}

% G ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )
% \renewcommand{\ottdruleAUNVarName}[0]{(Var$^{-[[≈au]]}$)}
% \renewcommand{\ottdruleAUShiftUName}[0]{($\uparrow^{[[≈au]]}$)}
% \renewcommand{\ottdruleAUForallName}[0]{($\forall^{[[≈au]]}$)}
% \renewcommand{\ottdruleAUArrowName}[0]{($\rightarrow^{[[≈au]]}$)}
% \renewcommand{\ottdruleAUAUName}[0]{(AU$^{-}$)}

\begin{lemma}[Completeness of Anti-Unification] \label{lemma:au-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            there exists $[[(Ξ', uQ', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$ 
                are anti-unification substitutions, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

            Then the anti-unification algorithm terminates, that is there exists
            $[[(Ξ, uQ, aus1, aus2)]]$ such that $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$

        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            there exists $[[(Ξ', uM', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$,
                are anti-unification substitutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

            Then the anti-unification algorithm succeeds, that is 
            there exists $[[(Ξ, uM, aus1, aus2)]]$ such that
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by the induction on $[[uM']]$ and mutually on $[[uQ']]$.
    \begin{caseof}
        \item $[[uM']] = [[â⁻]]$ 
            Then since $[[Γ ; · ⊢ aus'i : Ξ']]$,
            $[[Γ ⊢ [aus'i] uM']] = [[ iNi ]]$. 
            This way, \ruleref{\ottdruleAUAULabel} is always applicable
            if other rules are not.

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:au-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a algorithmic variable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:au-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:au-deterministic}).
    \end{itemize}

>>>>

\obsAuDeterministic*
\begin{proof}
    By trivial induction on $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
    and mutually on $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.
\end{proof}

\obsNamesDefinedByMapping*
\begin{proof}
    By simple induction on $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
    and mutually on $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$.
    Let us consider tha last rule applied to infer this judgment.
    \begin{caseof}
        \item \ruleref{\ottdruleAUPVarLabel} or \ruleref{\ottdruleAUNVarLabel},
        then $[[Ξ]] = [[·]]$, and the property holds vacuously.

@@

\lemmaAuSoundness*
\begin{proof}
    We prove it by induction on 
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
    and mutually, $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$.
    Let us consider the last rule applied to infer this judgement.
    \begin{caseof}
        \item \ruleref{\ottdruleAUNVarLabel}, then $[[iN1]] = [[α⁻]] = [[iN2]]$,
              $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, and $[[aus1]] = [[aus2]] = [[·]]$.
            \begin{enumerate}
                \item $[[Γ ; · ⊢ α⁻]]$ follows from the assumption $[[Γ ⊢ α⁻]]$,
                \item $[[Γ ; · ⊢ · : ·]]$ holds trivially, and
                \item $[[ [·] α⁻ = α⁻ ]]$ holds trivially.
            \end{enumerate}
        \item \label{case:anti-unification-soundness:shiftu}
         \ruleref{\ottdruleAUShiftULabel}, then $[[iN1]] = [[↑iP1]]$,
                $[[iN2]] = [[↑iP2]]$, and the algorithm makes the recursive call:
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
                returning $[[(Ξ, ↑uQ, aus1, aus2)]]$ as the result.

@@

\lemmaAuCompleteness*
\begin{proof}
    We prove it by the induction on $[[uM']]$ and mutually on $[[uQ']]$.
    \begin{caseof}
        \item $[[uM']] = [[â⁻]]$ 
            Then since $[[Γ ; · ⊢ aus'i : Ξ']]$,
            $[[Γ ⊢ [aus'i] uM']] = [[ iNi ]]$. 
            This way, \ruleref{\ottdruleAUAULabel} is always applicable
            if other rules are not.

@@

\lemmaAuInitial*
\begin{proof}
    First, let us assume that $[[uM']]$ is a algorithmic variable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:au-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:au-deterministic}).
    \end{itemize}

### 2023-11-13 proofs/tex/constraint-merge-lemmas.tex
Restructure proofs

\begin{observation}[Constraint Entry Merge is Deterministic]
    \label{obs:entry-merge-deterministic}
    For fixed $[[Γ]]$, $[[scE1]]$, $[[scE2]]$,
    if $[[Γ ⊢ scE1 & scE2 = scE]]$ and $[[Γ ⊢ scE1 & scE2 = scE']]$ 
    then $[[scE]] = [[scE']]$.
\end{observation}
\begin{proof}
    First, notice that the shape of $[[scE1]]$ and $[[scE2]]$
    uniquely determines the rule applied to infer  
    $[[Γ ⊢ scE1 & scE2 = scE]]$,
    which is consequently, the same rule used to 
    infer $[[Γ ⊢ scE1 & scE2 = scE']]$.
    Second, notice that the premises of each rule are deterministic
    on the input:
    the positive subtyping is deterministic by \cref{obs:pos-subt-deterministic},
    and the least upper bound is deterministic by \cref{obs:lub-deterministic}.
\end{proof}

\begin{observation}[Subgyping Constraint Merge is Deterministic]
    \label{obs:subt-merge-deterministic}
    Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$
    If $[[Θ ⊢ SC1 & SC2 = UC]]$ and $[[Θ ⊢ SC1 & SC2 = UC']]$ are defined then 
    $[[SC]] = [[SC']]$.
\end{observation}
\begin{proof}
    The proof is analogous to the proof of \cref{obs:unif-merge-deterministic} 
    but uses \cref{obs:entry-merge-deterministic} to show 
    that the merge of the matching constraint entries is fixed.
\end{proof}

\begin{lemma} [Soundness of Constraint Entry Merge]
    \label{lemma:entry-merge-soundness}
    For a fixed context $[[Γ]]$,
    suppose that  $[[Γ ⊢ scE1]]$ and $[[Γ ⊢ scE2]]$. 
    If $[[Γ ⊢ scE1 & scE2 = scE]]$ is defined then
    \begin{enumerate}
        \item $[[Γ ⊢ scE]]$
        \item For any $[[Γ ⊢ iP]]$, $[[Γ ⊢ iP : scE]]$ implies $[[Γ ⊢ iP : scE1]]$ and $[[Γ ⊢ iP : scE2]]$ 
    \end{enumerate}
\end{lemma}
\begin{proof}
    Let us consider the rule forming $[[Γ ⊢ scE1 & scE2 = scE]]$.
    \begin{caseof}
        \item \ruleref{\ottdruleSCMEPEqEqLabel}, i.e. 
            $[[Γ ⊢ scE1 & scE2 = scE]]$
            has form $[[Γ ⊢ (pua :≈ iQ) & (pua :≈ iQ') = (pua :≈ iQ)]]$
            and $[[nf(iQ) = nf(iQ')]]$. The latter implies $[[Γ ⊢ iQ ≈ iQ']]$ by
            \cref{lemma:subt-equiv-algorithmization}.
            Then
            \begin{enumerate}
                \item $[[Γ ⊢ scE]]$, i.e. $[[Γ ⊢ pua :≈ iQ]]$ holds by assumption;
                \item by inversion, $[[Γ ⊢ iP : (pua :≈ iQ)]]$ means $[[Γ ⊢ iP ≈ iQ]]$,
                and by transitivity of equivalence (\cref{corollary:equivalence-transitivity}), 
                $[[Γ ⊢ iP ≈ iQ']]$. Thus, $[[Γ ⊢ iP : scE1]]$ and $[[Γ ⊢ iP : scE2]]$ hold
                by \ruleref{\ottdruleSATSCEPEqLabel}.
            \end{enumerate}
        \item \ruleref{\ottdruleSCMENEqEqLabel} the negative case is proved in exactly the same way as the positive one.
        \item \ruleref{\ottdruleSCMESupSupLabel} 
            Then $[[scE1]]$ is $[[pua :≥ iQ1]]$, $[[scE2]]$ is $[[pua :≥ iQ2]]$,
            and $[[scE1 & scE2]] = [[scE]]$ is $[[pua :≥ iQ]]$ where $[[iQ]]$ is the least upper bound of $[[iQ1]]$ and $[[iQ2]]$.
            Then by \cref{lemma:lub-soundness},
            \begin{itemize}
                \item $[[Γ ⊢ iQ]]$,
                \item $[[Γ ⊢ iQ ≥ iQ1]]$,
                \item $[[Γ ⊢ iQ ≥ iQ2]]$.
            \end{itemize}

@@

\begin{lemma} [Soundness of Constraint Merge] \label{lemma:merge-soundness}
    Suppose that $[[Θ ⊢ SC1 : Ξ1]]$ and $[[Θ ⊢ SC2 : Ξ2]]$ 
    and $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
    Then 
    \begin{enumerate}
        \item $[[Θ ⊢ SC : Ξ1 ∪ Ξ2]]$,
        \item for any substitution $[[Θ ⊢ uσ : Ξ1 ∪ Ξ2]]$, 
            $[[ Θ ⊢ uσ : SC ]]$
            implies $[[ Θ ⊢ uσ : SC1 ]]$ and $[[ Θ ⊢ uσ : SC2 ]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    By definition, $[[Θ ⊢ SC1 & SC2 = SC]]$ consists of three parts:
    entries of $[[SC1]]$ that do not have matching entries of $[[SC2]]$,
    entries of $[[SC2]]$ that do not have matching entries of $[[SC1]]$,
    and the merge of matching entries.

@@

\begin{lemma} [Completeness of Constraint Entry Merge]
    \label{lemma:entry-merge-completeness}
    For a fixed context $[[Γ]]$,
    suppose that $[[Γ ⊢ scE1]]$ and $[[Γ ⊢ scE2]]$ are matching constraint entries.
    \begin{itemize}
        \item for a type $[[iP]]$ such that $[[Γ ⊢ iP : scE1]]$ and $[[Γ ⊢ iP : scE2]]$,
        $[[Γ ⊢ scE1 & scE2 = scE]]$ is defined and $[[Γ ⊢ iP : scE]]$.
        \item for a type $[[iN]]$ such that $[[Γ ⊢ iN : scE1]]$ and $[[Γ ⊢ iN : scE2]]$,
        $[[Γ ⊢ scE1 & scE2 = scE]]$ is defined and $[[Γ ⊢ iN : scE]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    Let us consider the shape of $[[scE1]]$ and $[[scE2]]$.
    \begin{caseof}
        \item $[[scE1]]$ is $[[pua :≈ iQ1]]$ and $[[scE2]]$ is $[[pua :≈ iQ2]]$.
            The proof repeats the corresponding case of \cref{lemma:unif-entry-merge-completeness}
        \item $[[scE1]]$ is $[[pua :≈ iQ1]]$ and $[[scE2]]$ is $[[pua :≥ iQ2]]$.
            Then $[[Γ ⊢ iP : scE1]]$ means $[[Γ ⊢ iP ≈ iQ1]]$, 
            and $[[Γ ⊢ iP : scE2]]$ means $[[Γ ⊢ iP ≥ iQ2]]$.
            Then by transitivity of subtyping, $[[Γ ⊢ iQ1 ≥ iQ2]]$,
            which means $[[Γ ; · ⊨ uQ1 ≥ iQ2 ⫤ ·]]$ by \cref{lemma:pos-subt-completeness}.
            This way, \ruleref{\ottdruleSCMEEqSupLabel} applies to infer
            $[[Γ ⊢ scE1 & scE2 = scE1]]$, and $[[Γ ⊢ iP : scE1]]$ holds by assumption.
        \item $[[scE1]]$ is $[[pua :≥ iQ1]]$ and $[[scE2]]$ is $[[pua :≥ iQ2]]$.
            Then $[[Γ ⊢ iP : scE1]]$ means $[[Γ ⊢ iP ≥ iQ1]]$, 
            and $[[Γ ⊢ iP : scE2]]$ means $[[Γ ⊢ iP ≥ iQ2]]$.
            By the completeness of the least upper bound (\cref{lemma:lub-completeness}), 
            $[[Γ ⊨ iQ1 ∨ iQ2 = iQ]]$, and $[[Γ ⊢ iP ≥ iQ]]$. 
            This way, \ruleref{\ottdruleSCMESupSupLabel} applies to infer
            $[[Γ ⊢ scE1 & scE2 = (pua :≥ iQ)]]$, 
            and $[[Γ ⊢ iP : (pua :≥ iQ)]]$ holds by \ruleref{\ottdruleSATSCESupLabel}.
        \item The negative cases are proved symmetrically.
    \end{caseof}
\end{proof}

\begin{lemma} [Completeness of Constraint Merge] 
    \label{lemma:merge-completeness}
    Suppose that $[[Θ ⊢ SC1 : Ξ1]]$ and $[[Θ ⊢ SC2 : Ξ2]]$.
    If there exists a substitution $[[Θ ⊢ uσ : Ξ1 ∪ Ξ2]]$ such that 
    $[[ Θ ⊢ uσ : SC1 ]]$ and $[[ Θ ⊢ uσ : SC2 ]]$
    then $[[Θ ⊢ SC1 & SC2 = SC]]$ is defined.
\end{lemma}
\begin{proof}
    By  definition, $[[SC1 & SC2]]$ is a union of
    \begin{enumerate}
        \item entries of $[[SC1]]$, which do not have matching entries in $[[SC2]]$,
        \item entries of $[[SC2]]$, which do not have matching entries in $[[SC1]]$, and 
        \item the merge of matching entries.
    \end{enumerate}

>>>>

\obsEntryMergeDeterministic*
\begin{proof}
    First, notice that the shape of $[[scE1]]$ and $[[scE2]]$
    uniquely determines the rule applied to infer  
    $[[Γ ⊢ scE1 & scE2 = scE]]$,
    which is consequently, the same rule used to 
    infer $[[Γ ⊢ scE1 & scE2 = scE']]$.
    Second, notice that the premises of each rule are deterministic
    on the input:
    the positive subtyping is deterministic by \cref{obs:pos-subt-deterministic},
    and the least upper bound is deterministic by \cref{obs:lub-deterministic}.
\end{proof}

\obsSubtMergeDeterministic*
\begin{proof}
    The proof is analogous to the proof of \cref{obs:unif-merge-deterministic} 
    but uses \cref{obs:entry-merge-deterministic} to show 
    that the merge of the matching constraint entries is fixed.
\end{proof}

\lemEntryMergeSoundness*
\begin{proof}
    Let us consider the rule forming $[[Γ ⊢ scE1 & scE2 = scE]]$.
    \begin{caseof}
        \item \ruleref{\ottdruleSCMEPEqEqLabel}, i.e. 
            $[[Γ ⊢ scE1 & scE2 = scE]]$
            has form $[[Γ ⊢ (pua :≈ iQ) & (pua :≈ iQ') = (pua :≈ iQ)]]$
            and $[[nf(iQ) = nf(iQ')]]$. The latter implies $[[Γ ⊢ iQ ≈ iQ']]$ by
            \cref{lemma:subt-equiv-algorithmization}.
            Then
            \begin{enumerate}
                \item $[[Γ ⊢ scE]]$, i.e. $[[Γ ⊢ pua :≈ iQ]]$ holds by assumption;
                \item by inversion, $[[Γ ⊢ iP : (pua :≈ iQ)]]$ means $[[Γ ⊢ iP ≈ iQ]]$,
                and by transitivity of equivalence (\cref{corollary:equivalence-transitivity}), 
                $[[Γ ⊢ iP ≈ iQ']]$. Thus, $[[Γ ⊢ iP : scE1]]$ and $[[Γ ⊢ iP : scE2]]$ hold
                by \ruleref{\ottdruleSATSCEPEqLabel}.
            \end{enumerate}
        \item \ruleref{\ottdruleSCMENEqEqLabel} the negative case is proved in exactly the same way as the positive one.
        \item \ruleref{\ottdruleSCMESupSupLabel} 
            Then $[[scE1]]$ is $[[pua :≥ iQ1]]$, $[[scE2]]$ is $[[pua :≥ iQ2]]$,
            and $[[scE1 & scE2]] = [[scE]]$ is $[[pua :≥ iQ]]$ where $[[iQ]]$ is the least upper bound of $[[iQ1]]$ and $[[iQ2]]$.
            Then by \cref{lemma:lub-soundness},
            \begin{itemize}
                \item $[[Γ ⊢ iQ]]$,
                \item $[[Γ ⊢ iQ ≥ iQ1]]$,
                \item $[[Γ ⊢ iQ ≥ iQ2]]$.
            \end{itemize}

@@

\lemMergeSoundness*
\begin{proof}
    By definition, $[[Θ ⊢ SC1 & SC2 = SC]]$ consists of three parts:
    entries of $[[SC1]]$ that do not have matching entries of $[[SC2]]$,
    entries of $[[SC2]]$ that do not have matching entries of $[[SC1]]$,
    and the merge of matching entries.

@@

\lemEntryMergeCompleteness*
\begin{proof}
    Let us consider the shape of $[[scE1]]$ and $[[scE2]]$.
    \begin{caseof}
        \item $[[scE1]]$ is $[[pua :≈ iQ1]]$ and $[[scE2]]$ is $[[pua :≈ iQ2]]$.
            The proof repeats the corresponding case of \cref{lemma:unif-entry-merge-completeness}
        \item $[[scE1]]$ is $[[pua :≈ iQ1]]$ and $[[scE2]]$ is $[[pua :≥ iQ2]]$.
            Then $[[Γ ⊢ iP : scE1]]$ means $[[Γ ⊢ iP ≈ iQ1]]$, 
            and $[[Γ ⊢ iP : scE2]]$ means $[[Γ ⊢ iP ≥ iQ2]]$.
            Then by transitivity of subtyping, $[[Γ ⊢ iQ1 ≥ iQ2]]$,
            which means $[[Γ ; · ⊨ uQ1 ≥ iQ2 ⫤ ·]]$ by \cref{lemma:pos-subt-completeness}.
            This way, \ruleref{\ottdruleSCMEEqSupLabel} applies to infer
            $[[Γ ⊢ scE1 & scE2 = scE1]]$, and $[[Γ ⊢ iP : scE1]]$ holds by assumption.
        \item $[[scE1]]$ is $[[pua :≥ iQ1]]$ and $[[scE2]]$ is $[[pua :≥ iQ2]]$.
            Then $[[Γ ⊢ iP : scE1]]$ means $[[Γ ⊢ iP ≥ iQ1]]$, 
            and $[[Γ ⊢ iP : scE2]]$ means $[[Γ ⊢ iP ≥ iQ2]]$.
            By the completeness of the least upper bound (\cref{lemma:lub-completeness}), 
            $[[Γ ⊨ iQ1 ∨ iQ2 = iQ]]$, and $[[Γ ⊢ iP ≥ iQ]]$. 
            This way, \ruleref{\ottdruleSCMESupSupLabel} applies to infer
            $[[Γ ⊢ scE1 & scE2 = (pua :≥ iQ)]]$, 
            and $[[Γ ⊢ iP : (pua :≥ iQ)]]$ holds by \ruleref{\ottdruleSATSCESupLabel}.
        \item The negative cases are proved symmetrically.
    \end{caseof}
\end{proof}

\lemMergeCompleteness*
\begin{proof}
    By  definition, $[[SC1 & SC2]]$ is a union of
    \begin{enumerate}
        \item entries of $[[SC1]]$, which do not have matching entries in $[[SC2]]$,
        \item entries of $[[SC2]]$, which do not have matching entries in $[[SC1]]$, and 
        \item the merge of matching entries.
    \end{enumerate}
